How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985

### 2022-06-09 Draft_June_2022/Full_draft_20220608.rmd
#weaved in policy object #added references in 2.1 and 3 #added references from environmental politics

---
title: How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article; names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "8 June 2022"
output:
  pdf_document: default
  fig_caption: yes
  html_document:
    df_print: paged
  word_document:
    reference_docx: template.docx
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: The Amazon Rainforest is one of the most important ecosystems in the world, and most of this biome is located in Brazil. Albeit the importance of understanding the relationship between the social construction of the Amazon and its effects on environmental outcomes, we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Discursive problem-construction within democracies legitimize ways of thinking and acting towards the environment while affecting decisions to deforest, or not, at the local level. Building on @hirschman1963 and @bacchi2009 conceptualization of chosen problems, we propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in political discourse. Combining hand-coding and supervised machine learning, we classify statements about the Amazon forest or region in 6130 official presidential speeches since 1985. We find that, first, international environmental events appear to drive presidents to speak more about the Amazon. Second, while the Amazon as a problem of economic integration dominated discourses from 1985 to the mid-2000s, environmental conservation and social development constructions briefly surpasses economic integration from 2010 to 2015. Constructing the Amazon as an issue of sovereignty also became more pertinent from 2010 onwards. Finally, using a multinomial regression model, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation, than economic integration or social development, as presidents move away from the Amazon region.
---

@@

We need to protect the Amazon from foreign interests. We need to exploit the Amazon’s natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem. Historically, the Brazilian government approached the Amazon region, forest and its people as a policy object.Different governmental programs contained an assumptions of what had to be solved in the Amazon, or in other words, it represents the Amazon as a particular problem. These specific problems touch on shared imaginaries of the Amazon as part of the larger socio-cultural history of the country as, for example, in relation to issues of national sovereignty, economic integration, social development, or environmental conservation.

@@

In this article, we investigate how the Brazilian Amazon has been constructed as a problem in political discourses. Building on @hirschman1963 concept of chosen problems and @bacchi2009 concept of problem-representation in policy, we propose a framework to identify problem-constructions in political discourses. Although problem-construction takes place in a series of instances, we analyze the case of discourses by Brazilian presidents since 1985. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify public policy, as well as shape its perception to broad audiences [@zarefsky2004]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations from governments’ response, generated from material and discursive governmental practices [@assuncao2015; @capobianco2019; @capobianco2021], are a crucial factor in decisions to deforest at the ground [@campbell2015]. Second, we argue that presidents can employ specific problem-constructions that build objects as specific problems depending on who they speak to. Presidential discourses take place in different sites, from launching a new bridge in a small municipality in the middle of the Amazon to a keynote speech in a business association in São Paulo or at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is socially constructed.

@@

Governments often have a choice in how they construct problems and what policy solutions they implement. In “Journey towards Progress”, @hirschman1963 draws a conceptual distinction between pressing problems (pressured from outside parties to the government) and chosen problems (chosen by the government at their own discretion). Pressing problems can be either privileged or neglected, depending on the degree of pressure exercised by the interested groups. Chosen problems are those governments select at their discretion. Problems can change from pressing to chosen across time and in space as a function of (a) solutions becoming available, (b) a change of level of government control in society, or (c) a shift of interests from top policymakers [@hirschman1963].

Additionally, @bacchi2009[p. 10] argues that policies have a cultural dimension as “it takes shape within specific historical and national or international contexts”. The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. The alleged problem is not always explicitly stated in policy: policies are represented as solutions for implicit problems.@hirschman1975[p. 388] exemplifies chosen problems with the case of the construction of Brasilia. Building Brasilia could solve a problem of regional inequality, a problem of a dormant economy without state investment, a problem of political representation, or all three. It is up to the discretion of the government to highlight or not the implicit problem a policy solves. A same policy object (e.g. Brasilia, the Amazon, criminality, poverty), thus, can be constructed as pressing or not depending on how a government represents it.

Governments can choose to emphasize (or not) one or more implicit problems that a policy solves depending on which interest group they are in communication with. Relative to how the policy is represented to be, it can be a solution of problems that are considered pressing or not for different groups and it is up to the discretion of the political actor to construct a particular problem in a particular way given context. @putnam1988 seminal article on the two-level game can help us make sense of this variation conceptually. The author argues that the outcomes of international negotiations lie within the overlap between the pressure of domestic groups and international groups. Taking the example of the 2012 Forest Code and the Amazon, President Dilma could emphasize its rural character for environmentalists at the Paris Summit, or the amnesty for deforestation for agricultural elites. The same policy, then, can solve the problem of environmental conservation as well as economic integration.

@@

An interesting site to study the relationship between varied problem-construction and policy issues is presidential speeches as they legitimize ways of thinking about an issue. While the relationship between presidential speeches and policy is not causal, they have the power to introduce and justify public policy, as well as shape its perception to broad audiences [@zarefsky2004]. Even if presidential discourse might not persuade public opinion regarding a proposed solution for a problem [@eshbaugh2010; @rottinghaus2009; @krebs2007], they direct the range of what is included in national debates [@druckman2004]. As presidents speak in varied locations to varied audiences, it is very likely that they chose to frame similar issues differently.

@@

Relatedly, the possibility of different problem-constructions has consequences for environmental outcomes as well. When Brazilian presidents speak about the Amazon it not only makes headlines, nationally and internationally [@brice2021; @harris2021; @miranda2021], but also incites responses, shapes expectations, and feeds into the behavior of many actors involved in the Amazon. Deforestation rates in Brazil are more responsive to the government’s environmental policy, generated from material and discursive governmental practices, than exogenous factors as market fluctuations [@assuncao2015; @capobianco2019; @capobianco2021]. Specifically, Capobianco finds that that the perception of higher risk of being caught for deforestation in the Amazon is the main mechanism through which environmental policy is effective. Relatedly, ethnographic research has also shown that expectations about governments’ response are a crucial factor in decisions to deforest at the ground [@campbell2015]. Thus, understanding diversity in presidential speeches is a first step to connect it to concrete environmental outcomes.

The connection between presidential discourse and the environment has been studied in the case of the United States but remains conceptually underdeveloped. @calderwood2019, for instance, examines 2919 mentions of climate change in American official presidential speeches since 1989. Among various findings, one that stands out is that American presidents frequently side-step the environmental aspects of climate change. He also identifies a shift from economic to security framing of climate issues, side-lining its environmental aspects. @calderwood2020 also tests the effect of geographic location and type of communication regarding climate change. Building prominently on @putnam1988, he hypothesizes that presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. The author finds evidence in support of his hypothesis, suggesting American presidential discourse at the top on global warming changes based on location. Elsewhere, @bevitori2015 investigates how the ‘environment’ has been constructed in American presidential discourse since 1960 using a more automated approach. The author finds that mentions of the environment are typically co-selected with the pronoun ‘our’, as well as with ‘economy’, ‘clean’, and ‘preserve’. These studies corroborate the possibilities of variation depending on the audience, but we are yet to tie vocabulary to specific problem-constructions based on wider shared meanings of environment and climate change. While these findings are US specific, they indicate that presidents raise different points about the environment at local, national, or international settings, depending on who they assume their audience is at that specific instance. Nevertheless, as @bacchi2009, we argue that problem-constructions touch on shared meanings about the region that are available to the speaker as part of larger social-cultural history. Mentions to the lexicon as ‘environment’, ‘preserve’, or ‘climate change’ should be tied to the larger meaning of policies in the United States and elsewhere.

On the other hand, there is also an empirical gap in terms of how the Amazon has been constructed as a problem along time and geographic location. While governmental discourses in Brazil have been studied for topics such as inflation or race relations, we only find one systematic analysis of Amazonian discourse. @barros2020 investigates Amazonian discourse in the Brazilian Congress with the objective of identifying the arguments put forth by congressmen. The main finding is that the economic value of the Amazon for the cattle industry is the most salient narrative, leading the author to conclude there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

@@

In *The Fate of the Forest: Developers, Destroyers, and Defenders of the Amazon*, @hecht1990[p. 1] write that all over the world tropical forests are destroyed, but “what imbues the case of the Amazon with such passion is the symbolic content of the dreams it ignites”. These dreams of fortunes to be found in the Amazon are composed of the imaginaries of bandeirantes from the southeast of Brazil and colonizers from everywhere else. It rendered the territory the venue for aspiration and the object of an intense scramble in the subsequent centuries, defined as “a (...) form of nation building (...)” [hecht1990, preface]. The Portuguese empire and subsequently the Brazilian monarchy were concerned with establishing their territory. In the process of securing Amazonian borders, Brazil thwarted “the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru” [@hecht2013, p. 8], and when the dust settled and the scramble was over, half of the Amazon emerged Brazilian. While Brazilian military diplomacy was very successful, the process did not come without its traumas. A significant experience was the negotiations with Bolivia in 1902 to secure the Amazonian state of Acre, during which they found out about American attempts to trick Brazil [@hecht1990]. This case was still part of the memory of the generals who led the country during the military dictatorship of 1964 and wanted to protect Brazil’s sovereignty over the Amazon from the communist threat during the Cold War [@garfield2013].

As we move from a world where non-state actors gain importance in environmental governance and international politics generally  [@silva-muller2022; @andonova2014; @westerwinter2021], the sovereignty problem becomes more varied. Multiple non-state actors (NGOs, foundations, IOs, and so on) join the conversation about Amazonian policies more substantially as the military dictatorship starts to end [@hochstetler2021; @capobianco2019; @franchini2019]. Threads to national sovereignty, consequently, can be interpreted as coming from a different set of actors than before. Allegedly false claims about the Brazilian Amazon in international and domestic fora, for instance, are often tied to strategies of ‘internationalizing’ the Amazon. This might come both from foreign actors as well as domestic non-state actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil while any foreign or non-state presence in the region is part of a broad strategy to take the region. The policy solutions to the issue of sovereignty, included the monitoring of the borders, strict regimes related to entry in the region, assertions of ownership nationally and internationally, and combating alleged disinformation from actors perceived as a threat.

@@

Closely related to economic integration, social development is a problem-construction that is often neglected while describing policy cycles or phases. While this might be related to social policies being more geared towards the whole nation, there are instances of specific social policies for the Amazon.Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples’ dignity, standards of living, and other constitutional rights when speaking about the Amazon. This is not only relevant as a construction, but speaks, minimally, to democratic concerns over providing electorates, especially in the Amazonian region, with better standards of living than previously.

@@

To analyze how presidents construct the Amazon as a problem in time, we build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil. Then, we proceed to identify all speeches that refer to the Amazon as a region, people, or forest in these speeches. We do so by detecting all speeches in which the stem “amazon” appears. In Portuguese, the stem captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

Using the poldis R package [@sposito2021], we proceed to extract two sentences before and two sentences after the sentence in which the stem “amazon” appears. We opt for picking two sentences before and two sentences after, rather than words, because sentences usually contain a cohesive idea. By doing so we create our unit of analysis: an Amazonian statement. We use Amazon statements as our unit of analysis for two main reasons. First, working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. Second, it increases the number of our observations and its variety meaningfully, allowing for more specificity in our analysis. This process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified. When an Amazonian statement contains two or more matches of the stem “amazon”, we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

We develop a codebook to code Amazonian statements in one or more problem-constructions (see codebook in appendix). In their conceptualization, each problem-construction is mutually exclusive, meaning that they cover different forms of constructing the Amazon as a problem. Nevertheless, each Amazonian statement might be assigned to one or more codes. A statement can, for example, construct the Amazon as a problem of sovereignty and a problem of economic integration, or a problem of social development and conservation. Amazonian statements, thus, can be either coded as pure-types or mixed-types. Mixed-types are relevant as constructions and policies are often multifaceted while portraying the Amazon as a combination of these problem-constructions. Figure 1, below, portrays this operationalization strategy.

\begin{landscape}

@@

\end{landscape}

With the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazon Statements identified. We chose to hand code half of the observations because since there are several nuances in discourse in how presidents talk about the amazon in time and as a problem, we conservatively code a relatively large training set for the subsequent automated machine learning. As well, this allows a robust validation set to verify the models. Still, automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. Intercoder agreement for each of the four main categories was 85%, on average. For each non-matching coded observation, the co-authors discussed and sorted their disagreements. The manually coded data is then randomly divided into a training set, containing 80% of the hand-coded observations (806 observations), and a validation set, containing the remaining 20% of the hand-coded data (201 observations).

We chose to employ a support-vector machine (SVM) algorithm to label texts, that is, a non- probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories [@meyer2021; @noble2006] . The SVM model is first trained using the hand coded training set and then employed to classify observations in the hand coded validation set. The trained SVM model was, on average 82%, accurate in classifying observations in the validation set before being tuned. After the SVM model is tuned, we use the model to automatically code the remaining 1007 Amazonian statements. The final dataset for analysis, excluding false positive matches, contains 1895 coded Amazonian statements. Finally, we extract locations for all speeches in the data. These locations represent the Brazilian state in which certain speech was given or an international country.

@@

This approach also comes with limitations. Our codebook is developed using specific Amazon-related vocabulary. For example, a statement will be coded as economic integration if it is meaningful support to the tax-free zone of Manaus or a Dam in the Amazon. However, the economy is generally a topic that presidents speak about. Hence, the high incidence of economic integration in Amazonian statements can also be related to the higher importance of this problem-construction in Brazil in time. Moreover, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: “amazon”. We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon “Amazon”, for example, when the president says, “the forest” or “deforestation”. Hence, we might be missing statements about Amazon that do not refer to it. However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Mata Atlantica. Nevertheless, our dataset covers only what is considered an official remark. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and more recently started to appear on social media. Problem- construction within presidential discourse, thus, also happens in different sites for which we do not account for in this paper.

@@

In 1990 the Amazon appeared in 12% of all speeches and, by 1992, this increased to 25% of all speeches. The 1992 Earth Summit, held in Brazil brought international attention to environmental topics in the Amazon. One of the big announcements, for example, was the consolidation of the first transnational partnership for the Amazon, the G7 Pilot Programme, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021]. Throughout the 1990s and 2000s, during the Cardoso and Lula administrations, the Amazon appeared, on average, on 15% of all speeches. Exceptionally, in 2009 we see another spike in the Amazon appearing in official speeches. This spike coincides with the 2009 Copenhagen Summit as well as the steepest decrease in deforestation rates since data is available. Lula led the delegation to Copenhagen with a self-image of “we do not promise, we deliver” when the stakes about climate change were high [@franchini2019].

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article; names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "8 June 2022"
output:
  html_document:
    df_print: paged
  fig_caption: yes
  pdf_document: default
  word_document:
    reference_docx: template.docx
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: The Amazon Rainforest is one of the most important ecosystems in the world, and most of this biome is located in Brazil. Albeit the importance of understanding the relationship between the social construction of the Amazon and its effects on environmental outcomes, we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Discursive problem-construction within democracies legitimize ways of thinking and acting towards the environment while affecting decisions to deforest, or not, at the local level. Building on @hirschman1963 and @bacchi2009 conceptualization of chosen problems, we propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in political discourse. Combining hand-coding and supervised machine learning, we classify statements about the Amazon forest or region in 6130 official presidential speeches since 1985. We find that, first, international environmental events appear to drive presidents to speak more about the Amazon. Second, while the Amazon as a problem of economic integration dominated discourses from 1985 to the mid-2000s, environmental conservation and social development constructions briefly surpasses economic integration from 2010 to 2015. Constructing the Amazon as an issue of sovereignty also became more pertinent from 2010 onwards. Finally, using a multinomial regression model, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation, than economic integration or social development, as presidents move away from the Amazon region.
---

@@

We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem. Historically, the Brazilian government approached the Amazon region, forest and its people as a policy object.Different governmental programs contained an assumptions of what had to be solved in the Amazon, or in other words, it represents the Amazon as a particular problem. These specific problems touch on shared imaginaries of the Amazon as part of the larger socio-cultural history of the country as, for example, in relation to issues of national sovereignty, economic integration, social development, or environmental conservation.

@@

In this article, we investigate how the Brazilian Amazon has been constructed as a problem in political discourses. Building on @hirschman1963 concept of chosen problems and @bacchi2009 concept of problem-representation in policy, we propose a framework to identify problem-constructions in political discourses. Although problem-construction takes place in a series of instances, we analyze the case of discourses by Brazilian presidents since 1985. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations from governments' response, generated from material and discursive governmental practices [@assuncao2015; @capobianco2019; @capobianco2021], are a crucial factor in decisions to deforest at the ground [@campbell2015]. Second, we argue that presidents can employ specific problem-constructions that build objects as specific problems depending on who they speak to. Presidential discourses take place in different sites, from launching a new bridge in a small municipality in the middle of the Amazon to a keynote speech in a business association in São Paulo or at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is socially constructed.

@@

Governments often have a choice in how they construct problems and what policy solutions they implement. In "Journey towards Progress", @hirschman1963 draws a conceptual distinction between pressing problems (pressured from outside parties to the government) and chosen problems (chosen by the government at their own discretion). Pressing problems can be either privileged or neglected, depending on the degree of pressure exercised by the interested groups. Chosen problems are those governments select at their discretion. Problems can change from pressing to chosen across time and in space as a function of (a) solutions becoming available, (b) a change of level of government control in society, or (c) a shift of interests from top policymakers [@hirschman1963].

Additionally, @bacchi2009 [p. 10] argues that policies have a cultural dimension as "it takes shape within specific historical and national or international contexts". The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. The alleged problem is not always explicitly stated in policy: policies are represented as solutions for implicit problems.\@hirschman1975[p. 388] exemplifies chosen problems with the case of the construction of Brasilia. Building Brasilia could solve a problem of regional inequality, a problem of a dormant economy without state investment, a problem of political representation, or all three. It is up to the discretion of the government to highlight or not the implicit problem a policy solves. A same policy object (e.g. Brasilia, the Amazon, criminality, poverty), thus, can be constructed as pressing or not depending on how a government represents it.

Governments can choose to emphasize (or not) one or more implicit problems that a policy solves depending on which interest group they are in communication with. Relative to how the policy is represented to be, it can be a solution of problems that are considered pressing or not for different groups and it is up to the discretion of the political actor to construct a particular problem in a particular way given context. @putnam1988 seminal article on the two-level game can help us make sense of this variation conceptually. The author argues that the outcomes of international negotiations lie within the overlap between the pressure of domestic groups and international groups. Relatedly, @hochstetler2012 argue that international commitments in the Brazilian case shift as a function of domestic cost-benefit calculations, as the country as a large emitter does not abide to the global commons framings. Taking the example of the 2012 Forest Code and the Amazon, President Dilma could emphasize its rural character for environmentalists at the Paris Summit, or the amnesty for deforestation for agricultural elites. The same policy, then, can solve the problem of environmental conservation as well as economic integration.

@@

An interesting site to study the relationship between varied problem-construction and policy issues is presidential speeches as they legitimize ways of thinking about an issue. While the relationship between presidential speeches and policy is not causal, they have the power to introduce and justify public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. Even if presidential discourse might not persuade public opinion regarding a proposed solution for a problem [@eshbaugh2010; @rottinghaus2009; @krebs2007], they direct the range of what is included in national debates [@druckman2004]. As presidents speak in varied locations to varied audiences, it is very likely that they chose to frame similar issues differently.

@@

Relatedly, the possibility of different problem-constructions has consequences for environmental outcomes as well. When Brazilian presidents speak about the Amazon it not only makes headlines, nationally and internationally [@brice2021; @harris2021; @miranda2021], but also incites responses, shapes expectations, and feeds into the behavior of many actors involved in the Amazon. Deforestation rates in Brazil are more responsive to the government's environmental policy, generated from material and discursive governmental practices, than exogenous factors as market fluctuations [@assuncao2015; @capobianco2019; @capobianco2021]. Specifically, Capobianco finds that that the perception of higher risk of being caught for deforestation in the Amazon is the main mechanism through which environmental policy is effective. Relatedly, ethnographic research has also shown that expectations about governments' response are a crucial factor in decisions to deforest at the ground [@campbell2015]. Thus, understanding diversity in presidential speeches is a first step to connect it to concrete environmental outcomes.

The connection between presidential discourse and the environment has been studied in the case of the United States but remains conceptually underdeveloped. @calderwood2019, for instance, examines 2919 mentions of climate change in American official presidential speeches since 1989. Among various findings, one that stands out is that American presidents frequently side-step the environmental aspects of climate change. He also identifies a shift from economic to security framing of climate issues, side-lining its environmental aspects. @calderwood2020 also tests the effect of geographic location and type of communication regarding climate change. Building prominently on @putnam1988, he hypothesizes that presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. The author finds evidence in support of his hypothesis, suggesting American presidential discourse at the top on global warming changes based on location. Elsewhere, @bevitori2015 investigates how the 'environment' has been constructed in American presidential discourse since 1960 using a more automated approach. The author finds that mentions of the environment are typically co-selected with the pronoun 'our', as well as with 'economy', 'clean', and 'preserve'. These studies corroborate the possibilities of variation depending on the audience, but we are yet to tie vocabulary to specific problem-constructions based on wider shared meanings of environment and climate change. While these findings are US specific, they indicate that presidents raise different points about the environment at local, national, or international settings, depending on who they assume their audience is at that specific instance. Nevertheless, as @bacchi2009, we argue that problem-constructions touch on shared meanings about the region that are available to the speaker as part of larger social-cultural history. Mentions to the lexicon as 'environment', 'preserve', or 'climate change' should be tied to the larger meaning of policies in the United States and elsewhere.

On the other hand, there is also an empirical gap in terms of how the Amazon has been constructed as a problem along time and geographic location. While governmental discourses in Brazil have been studied for topics such as inflation or race relations, we only find one systematic analysis of Amazonian discourse. @barros2020 investigates Amazonian discourse in the Brazilian Congress with the objective of identifying the arguments put forth by congressmen. The main finding is that the economic value of the Amazon for the cattle industry is the most salient narrative, leading the author to conclude there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development). We also find one analysis of Bolsonaro's rhetoric about the Amazon, but the study relies on a reduced sample without transparent explanations of how speeches were selected [@marquardt2022].

@@

In *The Fate of the Forest: Developers, Destroyers, and Defenders of the Amazon*, @hecht1990 [p. 1] write that all over the world tropical forests are destroyed, but "what imbues the case of the Amazon with such passion is the symbolic content of the dreams it ignites". These dreams of fortunes to be found in the Amazon are composed of the imaginaries of bandeirantes from the southeast of Brazil and colonizers from everywhere else. It rendered the territory the venue for aspiration and the object of an intense scramble in the subsequent centuries, defined as "a (...) form of nation building (...)" [hecht1990, preface]. The Portuguese empire and subsequently the Brazilian monarchy were concerned with establishing their territory. In the process of securing Amazonian borders, Brazil thwarted "the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru" [@hecht2013, p. 8], and when the dust settled and the scramble was over, half of the Amazon emerged Brazilian. While Brazilian military diplomacy was very successful, the process did not come without its traumas. A significant experience was the negotiations with Bolivia in 1902 to secure the Amazonian state of Acre, during which they found out about American attempts to trick Brazil [@hecht1990]. This case was still part of the memory of the generals who led the country during the military dictatorship of 1964 and wanted to protect Brazil's sovereignty over the Amazon from the communist threat during the Cold War [@garfield2013].

As we move from a world where non-state actors gain importance in environmental governance and international politics generally [@silva-muller2022; @andonova2014; @westerwinter2021], the sovereignty problem becomes more varied. Multiple non-state actors (NGOs, foundations, IOs, and so on) join the conversation about Amazonian policies more substantially as the military dictatorship starts to end [@hochstetler2021; @capobianco2019; @franchini2019]. Threads to national sovereignty, consequently, can be interpreted as coming from a different set of actors than before. Allegedly false claims about the Brazilian Amazon in international and domestic fora, for instance, are often tied to strategies of 'internationalizing' the Amazon. This might come both from foreign actors as well as domestic non-state actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil while any foreign or non-state presence in the region is part of a broad strategy to take the region. The policy solutions to the issue of sovereignty, included the monitoring of the borders, strict regimes related to entry in the region, assertions of ownership nationally and internationally, and combating alleged disinformation from actors perceived as a threat.

@@

Closely related to economic integration, social development is a problem-construction that is often neglected while describing policy cycles or phases. While this might be related to social policies being more geared towards the whole nation, there are instances of specific social policies for the Amazon. Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples' dignity, standards of living, and other constitutional rights when speaking about the Amazon. This is not only relevant as a construction, but speaks, minimally, to democratic concerns over providing electorates, especially in the Amazonian region, with better standards of living than previously.

@@

To analyze how presidents construct the Amazon as a problem in time, we build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil. Then, we proceed to identify all speeches that refer to the Amazon as a region, people, or forest in these speeches. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

Using the poldis R package [@sposito2021], we proceed to extract two sentences before and two sentences after the sentence in which the stem "amazon" appears. We opt for picking two sentences before and two sentences after, rather than words, because sentences usually contain a cohesive idea. By doing so we create our unit of analysis: an Amazonian statement. We use Amazon statements as our unit of analysis for two main reasons. First, working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. Second, it increases the number of our observations and its variety meaningfully, allowing for more specificity in our analysis. This process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified. When an Amazonian statement contains two or more matches of the stem "amazon", we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

We follow the procedures outlined by @grimmer2022 for supervised classification.First, we develop a codebook to code Amazonian statements in one or more problem-constructions (see codebook in appendix). In their conceptualization, each problem-construction is mutually exclusive, meaning that they cover different forms of constructing the Amazon as a problem. Nevertheless, each Amazonian statement might be assigned to one or more codes. A statement can, for example, construct the Amazon as a problem of sovereignty and a problem of economic integration, or a problem of social development and conservation. Amazonian statements, thus, can be either coded as pure-types or mixed-types. Mixed-types are relevant as constructions and policies are often multifaceted while portraying the Amazon as a combination of these problem-constructions. Figure 1, below, portrays this operationalization strategy.

@@

Second, with the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazon Statements identified. We chose to hand code half of the observations because since there are several nuances in discourse in how presidents talk about the amazon in time and as a problem, we conservatively code a relatively large training set for the subsequent automated machine learning. As well, this allows a robust validation set to verify the models. As the size of the training set should increase with the number of categories [@grimmer2022], we deemed four categories and 50% of the statements as reliable enough. Also, automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. Intercoder agreement for each of the four main categories was 85%, on average. For each non-matching coded observation, the co-authors discussed and sorted their disagreements.

Third, the hand-coded data is then randomly divided into a training set, containing 80% of the hand-coded observations (806 observations), and a validation set, containing the remaining 20% of the hand-coded data (201 observations). We chose to employ a support-vector machine (SVM) algorithm to label texts, that is, a non- probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories [@meyer2021; @noble2006] . The SVM model is first trained using the hand coded training set and then employed to classify observations in the hand coded validation set. The trained SVM model was, on average 82%, accurate in classifying observations in the validation set before being tuned. After the SVM model is tuned, we use the model to automatically code the remaining 1007 Amazonian statements. The final dataset for analysis, excluding false positive matches, contains 1895 coded Amazonian statements. Finally, we extract locations for all speeches in the data. These locations represent the Brazilian state in which certain speech was given or an international country.

@@

This approach also comes with limitations. Our codebook is developed using specific Amazon-related vocabulary. For example, a statement will be coded as economic integration if it is meaningful support to the tax-free zone of Manaus or a Dam in the Amazon. However, the economy is generally a topic that presidents speak about. Hence, the high incidence of economic integration in Amazonian statements can also be related to the higher importance of this problem-construction in Brazil in time. Moreover, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: "amazon". We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon "Amazon", for example, when the president says, "the forest" or "deforestation". Hence, we might be missing statements about Amazon that do not refer to it. However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Mata Atlantica. Nevertheless, our dataset covers only what is considered an official remark. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and more recently started to appear on social media. Problem- construction within presidential discourse, thus, also happens in different sites for which we do not account for in this paper.

@@

In 1990 the Amazon appeared in 12% of all speeches and, by 1992, this increased to 25% of all speeches. The 1992 Earth Summit, held in Brazil brought international attention to environmental topics in the Amazon. One of the big announcements, for example, was the consolidation of the first transnational partnership for the Amazon, the G7 Pilot Programme, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021]. Throughout the 1990s and 2000s, during the Cardoso and Lula administrations, the Amazon appeared, on average, on 15% of all speeches. Exceptionally, in 2009 we see another spike in the Amazon appearing in official speeches. This spike coincides with the 2009 Copenhagen Summit as well as the steepest decrease in deforestation rates since data is available. Lula led the delegation to Copenhagen with a self-image of "we do not promise, we deliver" when the stakes about climate change were high [@franchini2019].
